{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Models for detecting BHBH, BHNS and NSNS binaries with LISA

### 2021-11-16 paper/tex/sections/appendix.tex
update labels in appendix tables
proof read variations
most comments up to variations resolved

In this section we summarise the main assumptions and settings that we use when performing population synthesis for this work. For a more general overview of every setting see \citet{Broekgaarden+2021}.

@@

We simulate between 1 and 100 million massive binaries for each of 50 metallicities equally spaced in log space between $Z \in [0.0001, 0.022]$, where $Z$ is the mass fraction of heavy elements. We simulate more binaries for higher metallicities so that large enough sample of DCOs at each metallicity (since DCOs are formed at a lower rate at higher metallicities). These metallicities span the allowed metallicity range for the original fitting formulae on which COMPAS is based \citep{Hurley+2000}. This is repeated for \nMinusOneModels{} physics variations (see Section \ref{sec:variation_assumptions}) and so in total over two billion binaries were simulated.

Each binary is sampled from initial distributions for the primary and secondary masses as well as the separation. The primary mass, that is the mass of the initially more massive star, is restricted to $m_1 \in [5, 150] \unit{M_{\odot}}$, which spans the range of interest for NS and BH formation in binary systems, and drawn from the \citet{Kroupa+2001} initial mass function (IMF), $p(m_1) \propto m_1^{-2.3}$. The secondary mass, $m_2$, is drawn using the initial mass ratio of the binary, $q \equiv m_2 / m_1$, which we assume to be uniform on $[0, 1]$, therefore $p(q) = 1$ \citep[e.g.\ consistent with][]{Sana+2012}. We additionally restrict the secondary masses $m_2 \ge 0.1 \unit{M_{\odot}}$, which is approximately the minimum mass for a main sequence star. We assume that the initial separation follows a flat in the log distribution with $p(a_i) \propto 1 / a_i$ and $a_i \in [0.01, 1000] \unit{AU}$ \citep{Opik+1924, Abt+1983}. We assume that all binary orbits are circular at birth to reduce the dimensions of initial parameters. Since we focus on post-interaction binaries which will have circularised after mass transfer we argue this is an reasonable assumption (as many studies have in the past) and is likely not critical for predicting detection rates \citep{Hurley+2002, deMink+2015}.

We apply the adaptive importance sampling algorithm STROOPWAFEL \citep{Broekgaarden+2019} to improve the yield of our sample. This algorithm increases the prevalence of target DCOs (BHBHs, BHNSs and NSNSs in this case) in the sample and assigns each a weight, $w$, which represents the probability of drawing the DCO without STROOPWAFEL in effect.

\subsection{Physical assumptions in our fiducial model}\label{app:fiducial_physics}

@@

\textit{Wind mass loss:} We follow the wind prescription from \citet{Belczynski+2008}, which was based on results from Monte Carlo radiative transfer simulation of \citet{Vink+2000, Vink+2001}. We use the wind mass loss rates from \citet{Vink+2001} for stars above $12500 \unit{K}$ and the rates from \citet{Hurley+2000} for cooler stars. Additionally, we use a separate, higher wind mass loss rate for luminous blue variable (LBV) stars, following \citet{Belczynski+2008}, to mimic observed LBV eruptions for stars with luminosities and effective temperatures above the Humphreys-Davidson limit. We use the Wolf-Rayet-like mass loss rate from \citet{Hamann+1998} with an additional metallicity scaling from \citet{Vink+2005} for helium stars, and set $f_{\rm WR} = 1$. See \citet{COMPAS:2021methodsPaper}, Section 3 for the explicit equations.

\textit{Mass Transfer:} In determining the stability of mass transfer we use the $\zeta$-prescription, which compares the radial response of the star with the response of the Roche lobe radius to the mass transfer \citep[e.g.][]{Hjellming+1987}. The mass transfer efficiency, $\beta \equiv \Delta M_{\rm acc} / \Delta M_{\rm don}$, is defined as the fraction of the mass transferred by the donor that is actually accreted by the accretor. We limit the maximum accretion rate for stars to $\Delta M_{\rm acc} / \Delta t \le 10 M_{\rm acc} / \tau_{\rm KH}$, where $\tau_{\rm KH}$ is the Kelvin-Helmholtz timescale of the star \citep{Paczynski+1972, Hurley+2002}. The maximum accretion rate for compact objects is limited to the Eddington accretion rate. If more mass than these rates is accreted then we assume that the excess is lost through isotropic re-emission in the vicinity of the accreting star \citep[e.g.][]{Massevitch+1975, Soberman+1997}. We assume that all mass transfer from a stripped post-helium-burning-star (case BB) onto a neutron star or black hole is unstable \citep{Tauris+2015}.

\textit{Common-Envelope:} A common-envelope phase follows dynamically unstable mass transfer and we parameterise this using the $\alpha$-$\lambda$ prescription from \citet{Webbink+1984} and \citet{deKool+1990}. We assume $\alpha = 1$, such that all of the gravitational binding energy is available for the ejection of the envelope. For $\lambda$ we use the fitting formulae from \citet{Xu+2010, Xu+2010a}. We assume that any Hertzsprung gap donor stars that initiate a common-envelope phase will not survive this phase due to a lack of a steep density gradient between the core and envelope \citep{Taam+2000, Ivanova+2004, Klencki+2021}. This follows the `pessimistic' common-envelope scenario \citep[c.f.][]{Belczynski+2007}. We remove any binaries where the secondary immediately fills its Roche lobe upon the conclusion of the common-envelope phase as we treat these as failed common-envelope ejections, likely leading to a stellar merger.

\textit{Supernovae:} We draw the remnant masses and natal kick magnitudes from different distributions depending on the type of supernova that occurs. For stars undergoing a general core-collapse supernova, we use the \textit{delayed} supernova remnant mass prescription from \citet{Fryer+2012}. The \textit{delayed} prescription does not reproduce a neutron star black hole mass gap and we use this as our default as it has been shown to provide a better fit for observed populations of DCOs \citep[e.g.][]{Vigna-Gomez+2018}. We draw the natal kick magnitudes from a Maxwellian velocity distribution with a one-dimensional root-mean-square velocity dispersion of $\sigma_{\rm rms}^{\rm 1D} = 265 \unit{km}{s^{-1}}$ \citep{Lyne+1994, Hobbs+2005}. We assume that stars with helium core masses between $1.6$--$2.25 \unit{M_{\odot}}$ \citep{Hurley+2002} experience electron-capture supernovae (ECSN) \citep{Nomoto+1984, Nomoto+1987, Ivanova+2008}. We set all remnant masses to $1.26 \unit{M_{\odot}}$ in this case as an approximation of the solution to Equation 8 of \citet{Timmes+1996}. For these supernovae, we set $\sigma_{\rm rms}^{\rm 1D} = 30 \unit{km}{s^{-1}}$ \citep[e.g.][]{Pfahl+2002, Podsiadlowski+2004}. We assume that stars that undergo case BB mass transfer \citep{Dewi+2002} experience extreme stripping which leads to an ultra-stripped supernova \citep{Tauris+2013, Tauris+2015}. For these supernovae we calculate the remnant mass using the \citet{Fryer+2012} prescription and use $\sigma_{\rm rms}^{\rm 1D} = 30 \unit{km}{s^{-1}}$ (as with ECSN). Stars with final helium core masses between $35$-$135 \unit{M_{\odot}}$ are presumed to undergo a pair-instability, or pulsational pair-instability supernova \citep[e.g.][]{Woosley+2007, Farmer+2019}. We follow the prescription from \citet{Marchant+2019} as implemented in \citep{Stevenson+2019} for these supernovae. We assume that kicks are isotropic in the frame of the collapsing star. We adopt a maximum neutron star mass of $2.5 \unit{M_{\odot}}$ \citep[e.g.][]{Kalogera+1996, Fryer+2015, Margalit+2017} for the fiducial model and change the \citet{Fryer+2012} prescription accordingly.

\subsection{Model variations} \label{sec:variation_assumptions}
In addition to our fiducial model for the formation of DCOs, we explore \nMinusOneModels{} other models in which we change various aspects of the mass transfer, common-envelope, supernova and wind mass loss physics assumptions in order to assess the effect of their uncertainties on the overall double compact object detection rates and distributions. Each of the models varies a single physics assumption (fiducial assumptions are outlined in Section~\ref{app:fiducial_physics}) and these models are outlined in Table~\ref{tab:physics_variations}.

Our fiducial model is labelled model \modFid{}. Models \modRangeMT{} focus on changes to the mass transfer physics assumptions. We explore the effect of fixing the mass transfer efficiency $\beta$ to a constant value, rather than allowing it to vary based on the maximum accretion rate. In models \modBetaLow{}, \modBetaMed{}, \modBetaHigh{}, in which we set the value of $\beta$ to $0.25$, $0.5$ and $0.75$ respectively. In model \modCaseBB{} we investigate the consequence of assuming that case BB mass transfer onto a neutron star or black hole is always stable rather than always unstable.

Models \modRangeCE{} focus on altering the common-envelope physics. We change the common-envelope efficiency parameter to $\alpha_{\rm CE} = 0.1, 0.5, 2.0, 10.0$ in models \modAlphaLowest{}, \modAlphaLow{}, \modAlphaHigh{} and \modAlphaHighest{} respectively. In model \modOpt{}, we relax our restriction that Hertzsprung gap donor stars cannot survive common-envelope events, thereby following the `optimistic' common-envelope scenario. We combine this with model \modCaseBB{} in model \modCaseBBOpt{}.

In models \modRangeSN{} we consider changes related to our assumptions about supernova physics. Model \modRapid{} uses the alternate \textit{rapid} remnant mass prescription from \citet{Fryer+2012} instead of the \textit{delayed} prescription. We change the maximum neutron star mass in models \modNSLow{} and \modNSHigh{} to $2$ and $3 \unit{M_{\odot}}$ respectively to account for the range of predicted maximum neutron star masses. Model \modNoPISN{} removes the implementation of pair-instability and pulsational pair-instability supernovae. In models \modSigLow{} and \modSigLower{} we decrease the root-mean-square velocity dispersion for core-collapse supernovae to explore the effect of lower kicks. Model \modNoBH{} removes the natal kick for all black holes.

Finally, in models \modRangeML{} we investigate the effect of changing our assumption about wind mass loss rates, specifically for Wolf-Rayet winds. We vary $f_{\rm WR}$ to $0.1$ and $5.0$ in models \modWRLow{} and \modWRHigh{} respectively. These values approximately span the current range of possible Wolf-Rayet wind efficiencies suggested from observations \citep[e.g.][]{Vink+2017, Hamann+2019, Shenar+2019, Miller-Jones+2021, vanSon+2021}.

\begin{table}[htb]
    \centering
    \begin{tabular}{cl}
        \hline \hline
        Model & Physics Variation \\
        \hline \hline
        \modFid & Fiducial (see Section~\ref{app:fiducial_physics}) \\
        \hline
        \modBetaLow & Fixed mass transfer efficiency of $\beta=0.25$ \\ 
        \modBetaMed & Fixed mass transfer efficiency of $\beta=0.5$  \\ 
        \modBetaHigh & Fixed mass transfer efficiency of $\beta=0.75$ \\ 
        \modCaseBB & Case BB mass transfer is always unstable \\
        \modCaseBBOpt & Model \modCaseBB{} + Model \modOpt{} \\
        \hline
        \modAlphaLowest & CE efficiency parameter $\alpha = 0.1$ \\
        \modAlphaLow & CE efficiency parameter $\alpha = 0.5$ \\
        \modAlphaHigh & CE efficiency parameter $\alpha = 2$   \\
        \modAlphaHighest & CE efficiency parameter $\alpha = 10$   \\
        \modOpt & HG donor stars initiating a CE survive CE \\
        \hline
        \modRapid & Fryer rapid SN remnant mass prescription \\
        \modNSLow & Maximum NS mass is fixed to $2\unit{M_{\rm \odot}}$ \\
        \modNSHigh & Maximum NS mass is fixed to $3\unit{M_{\rm \odot}}$ \\
        \modNoPISN & PISN and pulsational-PISN not implemented \\
        \modSigLow & $\sigma_{\rm{rms}}^{\rm{1D}}=100 \unit{km}{s^{-1}}$ for core-collapse supernova \\  
        \modSigLower & $\sigma_{\rm{rms}}^{\rm{1D}}=30  \unit{km}{s^{-1}}$ for core-collapse supernova \\ 
        \modNoBH & Black holes receive no natal kick \\
        \hline
        \modWRLow & Wolf-Rayet wind factor $f_{\rm WR} = 0.1$ \\
        \modWRHigh & Wolf-Rayet wind factor $f_{\rm WR} = 5.0$ \\
        \hline \hline
    \end{tabular}%
    \caption{A description of the \nModels{} binary population synthesis models used in this study. \modFid{} is the fiducial model, \modRangeMT{} change mass transfer physics, \modRangeCE{} change common-envelope physics , \modRangeSN{} change supernova physics and \modRangeML{} change wind mass loss \citep[c.f.][Table 2]{Broekgaarden+2021}.}
    \label{tab:physics_variations}
\end{table}

@@

Moreover, the distribution of parameters within the population, particularly the mass distributions, are notably disparate. By using only two fixed metallicity populations, unphysical artifacts are introduced into distribution of DCO masses (Kummer et al. (in prep)). For example, in Fig.~\ref{fig:bh_mass_simple_mw}, we show the black hole mass distribution produced by the simulation using the simple Milky Way model. Despite the fact that these KDEs use the same bandwidth as Fig.~\ref{fig:fiducial_pdf_distributions}, the distributions show many more sharp transitions, which is a result of pileups occurring at specific masses for specific metallicities. Moreover, the lack of lower metallicities systems means that higher mass systems are not formed and so we see the distributions do not include a high mass tail such as in our fiducial results.

@@

    \caption{The number of detectable binaries in a 4- and 10-year LISA mission for the \nModels{} different model variations and each DCO type. The `All' column contains the total expected detections when summed over the three types. The final two rows show the minimum and maximum rates across all model variations. We embolden the corresponding rate for convenience of seeing which variation results in the minimum/maximum. Each value shows the mean and the 1-$\sigma$ Poisson uncertainty. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}

@@

        E & unstable
        case BB & \confinv{76.7}{8.7}{8.3} & \confinv{69.3}{8.3}{8.7} & \confinv{7.3}{2.3}{2.7} & \boldconfinv{0.2}{0.2}{0.8} & \confinv{121.4}{11.4}{10.6} & \confinv{109.3}{10.3}{10.7} & \confinv{11.8}{3.8}{3.2} & \boldconfinv{0.4}{0.4}{0.6}\\
        F & E+K & \boldconfinv{320.9}{17.9}{18.1} & \boldconfinv{154.3}{12.3}{12.7} & \boldconfinv{148.4}{12.4}{12.6} & \confinv{18.2}{4.2}{3.8} & \boldconfinv{483.7}{21.7}{22.3} & \boldconfinv{239.5}{15.5}{15.5} & \confinv{216.8}{14.8}{15.2} & \confinv{27.4}{5.4}{5.6}\\

@@

\begin{figure}[p]
    \centering
    \includegraphics[height=0.85\textheight]{fig16_formation_channels.pdf}
    \caption{Fraction of each DCO type that is formed through different formation channels for all physics variations. Channels are described in detail in \citet{Broekgaarden+2021}. The classic, single core CEE and double core CEE channels all require at least one common-envelope event whilst only `only stable' consists of only stable mass transfer and `other' contains the remaining binaries which are mainly formed from `lucky' supernova kicks that shrink the binary. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig16_formation_channels.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/formation_channels.ipynb}{\faBook}.}
    \label{fig:formation_channels}
\end{figure}

@@

    \includegraphics[width=0.8\textwidth]{fig18_dco_formation_distributions.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the properties of the detectable systems at DCO formation. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig18_dco_formation_distributions.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:dco_formation_properties}

    \includegraphics[width=\textwidth]{fig19_dco_relative_rates.pdf}
    \caption{Similar to Fig.~\ref{fig:detection_rates}, this shows the \textit{relative} detection rates compared to the fiducial model in order to emphasise how much each DCO is affected by model variations compared to other DCOs. The scatter points show the mean and the shaded areas show the 1-$\sigma$ uncertainty range. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig19_dco_relative_rates.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:dco_relative_rates}
\end{figure*}

>>>>

In this section we summarise the main assumptions and settings from the population synthesis simulation from Broekgaarden et al.\ (2021a, 2021b).

@@

Broekgaarden et al.\ (2021a, 2021b) simulate between 1 and 100 million massive binaries for each of 50 metallicities equally spaced in log space between $Z \in [0.0001, 0.022]$, where $Z$ is the mass fraction of heavy elements. They simulate more binaries for higher metallicities so that large enough sample of DCOs at each metallicity (since DCOs are formed at a lower rate at higher metallicities). These metallicities span the allowed metallicity range for the original fitting formulae on which COMPAS is based \citep{Hurley+2000}. This is repeated for \nMinusOneModels{} physics variations (see Section \ref{sec:variation_assumptions}) and so in total over two billion binaries were simulated.

Each binary is sampled from initial distributions for the primary and secondary masses as well as the separation. The primary mass, that is the mass of the initially more massive star, is restricted to $m_1 \in [5, 150] \unit{M_{\odot}}$, which spans the range of interest for NS and BH formation in binary systems, and drawn from the \citet{Kroupa+2001} initial mass function (IMF), $p(m_1) \propto m_1^{-2.3}$. The secondary mass, $m_2$, is drawn using the initial mass ratio of the binary, $q \equiv m_2 / m_1$, which Broekgaarden et al.\ (2021a, 2021b) assume to be uniform on $[0, 1]$, therefore $p(q) = 1$ \citep[e.g.\ consistent with][]{Sana+2012}. They additionally restrict the secondary masses $m_2 \ge 0.1 \unit{M_{\odot}}$, which is approximately the minimum mass for a main sequence star. They assume that the initial separation follows a flat in the log distribution with $p(a_i) \propto 1 / a_i$ and $a_i \in [0.01, 1000] \unit{AU}$ \citep{Opik+1924, Abt+1983}. They assume that all binary orbits are circular at birth to reduce the dimensions of initial parameters. Since they focus on post-interaction binaries which will have circularised after mass transfer they argue this is an reasonable assumption (as many studies have in the past) and is likely not critical for predicting detection rates \citep{Hurley+2002, deMink+2015}.

Broekgaarden et al.\ (2021a, 2021b) apply the adaptive importance sampling algorithm STROOPWAFEL \citep{Broekgaarden+2019} to improve the yield of their sample. This algorithm increases the prevalence of target DCOs (BHBHs, BHNSs and NSNSs in this case) in the sample and assigns each a weight, $w$, which represents the probability of drawing the DCO without STROOPWAFEL in effect.

\subsection{Physical assumptions in the fiducial model}\label{app:fiducial_physics}

@@

\textit{Wind mass loss:} Broekgaarden et al.\ (2021a, 2021b) follow the wind prescription from \citet{Belczynski+2008}, which was based on results from Monte Carlo radiative transfer simulation of \citet{Vink+2000, Vink+2001}. They use the wind mass loss rates from \citet{Vink+2001} for stars above $12500 \unit{K}$ and the rates from \citet{Hurley+2000} for cooler stars. Additionally, they use a separate, higher wind mass loss rate for luminous blue variable (LBV) stars, following \citet{Belczynski+2008}, to mimic observed LBV eruptions for stars with luminosities and effective temperatures above the Humphreys-Davidson limit. They use the Wolf-Rayet-like mass loss rate from \citet{Hamann+1998} with an additional metallicity scaling from \citet{Vink+2005} for helium stars, and set $f_{\rm WR} = 1$. See \citet{COMPAS:2021methodsPaper}, Section 3 for the explicit equations.

\textit{Mass Transfer:} In determining the stability of mass transfer Broekgaarden et al.\ (2021a, 2021b) use the $\zeta$-prescription, which compares the radial response of the star with the response of the Roche lobe radius to the mass transfer \citep[e.g.][]{Hjellming+1987}. The mass transfer efficiency, $\beta \equiv \Delta M_{\rm acc} / \Delta M_{\rm don}$, is the fraction of the mass transferred by the donor that is actually accreted by the accretor. They limit the maximum accretion rate for stars to $\Delta M_{\rm acc} / \Delta t \le 10 M_{\rm acc} / \tau_{\rm KH}$, where $\tau_{\rm KH}$ is the Kelvin-Helmholtz timescale of the star \citep{Paczynski+1972, Hurley+2002}. The maximum accretion rate for compact objects is limited to the Eddington accretion rate. If more mass than these rates is accreted then they assume that the excess is lost through isotropic re-emission in the vicinity of the accreting star \citep[e.g.][]{Massevitch+1975, Soberman+1997}. They assume that all mass transfer from a stripped post-helium-burning-star (case BB) onto a neutron star or black hole is unstable \citep{Tauris+2015}.

\textit{Common-Envelope:} A common-envelope phase follows dynamically unstable mass transfer and Broekgaarden et al.\ (2021a, 2021b) parameterise this using the $\alpha$-$\lambda$ prescription from \citet{Webbink+1984} and \citet{deKool+1990}. They assume $\alpha = 1$, such that all of the gravitational binding energy is available for the ejection of the envelope. For $\lambda$ they use the fitting formulae from \citet{Xu+2010, Xu+2010a}. They assume that any Hertzsprung gap donor stars that initiate a common-envelope phase will not survive this phase due to a lack of a steep density gradient between the core and envelope \citep{Taam+2000, Ivanova+2004, Klencki+2021}. This follows the `pessimistic' common-envelope scenario \citep[c.f.][]{Belczynski+2007}. They remove any binaries where the secondary immediately fills its Roche lobe upon the conclusion of the common-envelope phase as they treat these as failed common-envelope ejections, likely leading to a stellar merger.

\textit{Supernovae:} Broekgaarden et al.\ (2021a, 2021b) draw the remnant masses and natal kick magnitudes from different distributions depending on the type of supernova that occurs. For stars undergoing a general core-collapse supernova, they use the \textit{delayed} supernova remnant mass prescription from \citet{Fryer+2012}. The \textit{delayed} prescription does not reproduce a neutron star black hole mass gap and they use this as their default as it has been shown to provide a better fit for observed populations of DCOs \citep[e.g.][]{Vigna-Gomez+2018}. They draw the natal kick magnitudes from a Maxwellian velocity distribution with a one-dimensional root-mean-square velocity dispersion of $\sigma_{\rm rms}^{\rm 1D} = 265 \unit{km}{s^{-1}}$ \citep{Lyne+1994, Hobbs+2005}. They assume that stars with helium core masses between $1.6$--$2.25 \unit{M_{\odot}}$ \citep{Hurley+2002} experience electron-capture supernovae (ECSN) \citep{Nomoto+1984, Nomoto+1987, Ivanova+2008}. They set all remnant masses to $1.26 \unit{M_{\odot}}$ in this case as an approximation of the solution to Equation 8 of \citet{Timmes+1996}. For these supernovae, they set $\sigma_{\rm rms}^{\rm 1D} = 30 \unit{km}{s^{-1}}$ \citep[e.g.][]{Pfahl+2002, Podsiadlowski+2004}. They assume that stars that undergo case BB mass transfer \citep{Dewi+2002} experience extreme stripping which leads to an ultra-stripped supernova \citep{Tauris+2013, Tauris+2015}. For these supernovae they calculate the remnant mass using the \citet{Fryer+2012} prescription and use $\sigma_{\rm rms}^{\rm 1D} = 30 \unit{km}{s^{-1}}$ (as with ECSN). Stars with final helium core masses between $35$-$135 \unit{M_{\odot}}$ are presumed to undergo a pair-instability, or pulsational pair-instability supernova \citep[e.g.][]{Woosley+2007, Farmer+2019}. They follow the prescription from \citet{Marchant+2019} as implemented in \citep{Stevenson+2019} for these supernovae. They assume that kicks are isotropic in the frame of the collapsing star. They adopt a maximum neutron star mass of $2.5 \unit{M_{\odot}}$ \citep[e.g.][]{Kalogera+1996, Fryer+2015, Margalit+2017} for the fiducial model and change the \citet{Fryer+2012} prescription accordingly.

\subsection{Model variations} \label{sec:variation_assumptions}
In addition to their fiducial model for the formation of DCOs, Broekgaarden et al.\ (2021a, 2021b) explore \nMinusOneModels{} other models in which they change various aspects of the mass transfer, common-envelope, supernova and wind mass loss physics assumptions in order to assess the effect of their uncertainties on the overall double compact object detection rates and distributions. Each of the models varies a single physics assumption (fiducial assumptions are outlined in Section~\ref{app:fiducial_physics}) and these models are outlined in Table~\ref{tab:physics_variations}.

Their fiducial model is labelled model \modFid{}. Models \modRangeMT{} focus on changes to the mass transfer physics assumptions. They explore the effect of fixing the mass transfer efficiency $\beta$ to a constant value, rather than allowing it to vary based on the maximum accretion rate. In models \modBetaLow{}, \modBetaMed{}, \modBetaHigh{}, in which they set the value of $\beta$ to $0.25$, $0.5$ and $0.75$ respectively.

Models \modRangeCE{} focus on altering the common-envelope physics.  In model \modCaseBB{} Broekgaarden et al.\ (2021a, 2021b) investigate the consequence of assuming that case BB mass transfer onto a neutron star or black hole always leads to a stellar merger. They change the common-envelope efficiency parameter to $\alpha_{\rm CE} = 0.1, 0.5, 2.0, 10.0$ in models \modAlphaLowest{}, \modAlphaLow{}, \modAlphaHigh{} and \modAlphaHighest{} respectively. In model \modOpt{}, they relax their restriction that Hertzsprung gap donor stars cannot survive common-envelope events, thereby following the `optimistic' common-envelope scenario. They combine this with model \modCaseBB{} in model \modCaseBBOpt{}.

In models \modRangeSN{} they consider changes related to their assumptions about supernova physics. Model \modRapid{} uses the alternate \textit{rapid} remnant mass prescription from \citet{Fryer+2012} instead of the \textit{delayed} prescription. They change the maximum neutron star mass in models \modNSLow{} and \modNSHigh{} to $2$ and $3 \unit{M_{\odot}}$ respectively to account for the range of predicted maximum neutron star masses. Model \modNoPISN{} removes the implementation of pair-instability and pulsational pair-instability supernovae. In models \modSigLow{} and \modSigLower{} they decrease the root-mean-square velocity dispersion for core-collapse supernovae to explore the effect of lower kicks. Model \modNoBH{} removes the natal kick for all black holes.

Finally, in models \modRangeML{} Broekgaarden et al.\ (2021a, 2021b) investigate the effect of changing their assumption about wind mass loss rates, specifically for Wolf-Rayet winds. They vary $f_{\rm WR}$ to $0.1$ and $5.0$ in models \modWRLow{} and \modWRHigh{} respectively. These values approximately span the current range of possible Wolf-Rayet wind efficiencies suggested from observations \citep[e.g.][]{Vink+2017, Hamann+2019, Shenar+2019, Miller-Jones+2021, vanSon+2021}.

\begin{table}[htb]
    \centering
    \begin{tabular}{cl}
        \hline \hline
        Model & Physics Variation \\
        \hline \hline
        \modFid & Fiducial (see Section~\ref{app:fiducial_physics}) \\
        \hline
        \modBetaLow & Fixed mass transfer efficiency of $\beta=0.25$ \\ 
        \modBetaMed & Fixed mass transfer efficiency of $\beta=0.5$  \\ 
        \modBetaHigh & Fixed mass transfer efficiency of $\beta=0.75$ \\
        \hline
        \modCaseBB & Exclude Case BB mass transfer \\
        \modCaseBBOpt & Case BB always unstable + Optimistic CE \\
        \modAlphaLowest & CE efficiency parameter $\alpha = 0.1$ \\
        \modAlphaLow & CE efficiency parameter $\alpha = 0.5$ \\
        \modAlphaHigh & CE efficiency parameter $\alpha = 2$   \\
        \modAlphaHighest & CE efficiency parameter $\alpha = 10$   \\
        \modOpt & HG donor stars initiating a CE survive CE \\
        \hline
        \modRapid & Fryer rapid SN remnant mass prescription \\
        \modNSLow & Maximum NS mass is fixed to $2\unit{M_{\rm \odot}}$ \\
        \modNSHigh & Maximum NS mass is fixed to $3\unit{M_{\rm \odot}}$ \\
        \modNoPISN & PISN and pulsational-PISN not implemented \\
        \modSigLow & $\sigma_{\rm{rms}}^{\rm{1D}}=100 \unit{km}{s^{-1}}$ for core-collapse supernova \\  
        \modSigLower & $\sigma_{\rm{rms}}^{\rm{1D}}=30  \unit{km}{s^{-1}}$ for core-collapse supernova \\ 
        \modNoBH & Black holes receive no natal kick \\
        \hline
        \modWRLow & Wolf-Rayet wind factor $f_{\rm WR} = 0.1$ \\
        \modWRHigh & Wolf-Rayet wind factor $f_{\rm WR} = 5.0$ \\
        \hline \hline
    \end{tabular}%
    \caption{A description of the \nModels{} binary population synthesis models used in this study. \modFid{} is the fiducial model, \modRangeMT{} change mass transfer physics, \modRangeCE{} change common-envelope physics , \modRangeSN{} change supernova physics and \modRangeML{} change wind mass loss \citep[c.f.][Table 2]{Broekgaarden+2021}.}
    \label{tab:physics_variations}
\end{table}

@@

Moreover, the distribution of parameters within the population, particularly the mass distributions, are notably disparate. By using only two fixed metallicity populations, unphysical artifacts are introduced into distribution of DCO masses (e.g.\ \citealt{Dominik+2015, Neijssel+2019}, Kummer et al. 2021). For example, in Fig.~\ref{fig:bh_mass_simple_mw}, we show the black hole mass distribution produced by the simulation using the simple Milky Way model. Despite the fact that these KDEs use the same bandwidth as Fig.~\ref{fig:fiducial_pdf_distributions}, the distributions show many more sharp transitions, which is a result of pileups occurring at specific masses for specific metallicities. Moreover, the lack of lower metallicities systems means that higher mass systems are not formed and so we see the distributions do not include a high mass tail such as in our fiducial results.

@@

    \caption{The number of detectable binaries in a 4- and 10-year LISA mission for the \nModels{} different model variations and each DCO type. Each model variation is discussed in App.~\ref{sec:variation_assumptions} and the trends in detection rates are discussed in Sec.~\ref{sec:detection_rate_analysis}. The `All' column contains the total expected detections when summed over the three types. The final two rows show the minimum and maximum rates across all model variations. We embolden the corresponding rate for convenience of seeing which variation results in the minimum/maximum. Each value shows the mean and the 1-$\sigma$ Poisson uncertainty. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}

@@

        E & no case BB & \confinv{76.7}{8.7}{8.3} & \confinv{69.3}{8.3}{8.7} & \confinv{7.3}{2.3}{2.7} & \boldconfinv{0.2}{0.2}{0.8} & \confinv{121.4}{11.4}{10.6} & \confinv{109.3}{10.3}{10.7} & \confinv{11.8}{3.8}{3.2} & \boldconfinv{0.4}{0.4}{0.6}\\
        F & E + K & \boldconfinv{320.9}{17.9}{18.1} & \boldconfinv{154.3}{12.3}{12.7} & \boldconfinv{148.4}{12.4}{12.6} & \confinv{18.2}{4.2}{3.8} & \boldconfinv{483.7}{21.7}{22.3} & \boldconfinv{239.5}{15.5}{15.5} & \confinv{216.8}{14.8}{15.2} & \confinv{27.4}{5.4}{5.6}\\

@@

\begin{figure}[p]
    \centering
    \includegraphics[height=0.85\textheight]{fig16_formation_channels.pdf}
    \caption{Fraction of each DCO type that is formed through different formation channels for all physics variations. Channels are described in detail in \citet{Broekgaarden+2021}. The classic, single core CEE and double core CEE channels all require at least one common-envelope event whilst only ``only stable'' consists of only stable mass transfer and ``other'' contains the remaining binaries which are mainly formed from case A ``classic'' binaries as well as ``lucky'' supernova kicks that shrink the binary. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig16_formation_channels.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/formation_channels.ipynb}{\faBook}.}
    \label{fig:formation_channels}
\end{figure}

@@

    \includegraphics[width=0.8\textwidth]{fig18_dco_formation_distributions.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the properties of the detectable systems at DCO formation. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig18_dco_formation_distributions.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:dco_formation_properties}
\end{figure*}
